---
title: "Life Expectancy Around The World"
author: "Grant Branstetter, Kayla Tounalom, Ysabelle Lara Olairez, Lizeth Velderrain-Lopez"
date: "2/16/2022"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
```{r}
source("../source/Education_Chart.R")
```


### Summary Information


### Summary Table 


### Chart 1
```{r pressure, echo=FALSE}
plot(education_plot)
```


We chose to use a scatterplot to compare the number of schooling years and life expectancy across the world to showcase the correlation whether it be a negative or positive correlation. As a result of the scatterplot, you can see the x-axis being schooling and y-axis being life expectancy, and the blue dots representing the data. As one can see in the chart there is a high positive correlation between life expectancy and years of schooling throughout the world. For example for years of schooling above 10 years, the average life expectancy of a person is higher compared to years of schooling for 5 years. For each year add of schooling and gaining an education the life expectancy increases. 

### Chart 2 

### Chart 3 


### Chart 4
